---
title: "random Forest"
author: "Angel Chen"
date: "12/7/2018"
output: pdf_document
---

```{r}
library(tidyverse)

```

```{r, message = FALSE, warning=FALSE}
setwd("~/Downloads")
dat_raw = read.csv("data.csv")
dat_raw = dat_raw[, 1:32]
```

```{r}
dat_covar = dat_raw %>% select(-id, -diagnosis)
dat_label = cbind(dat_raw$diagnosis) %>% as.data.frame
```



```{r}
set.seed(123)
index = sample(1:nrow(dat_raw), size = trunc(.8 * nrow(dat_raw)))
dat_raw$diagnosis = as.integer(factor(dat_raw$diagnosis))-1
Train <- dat_raw[index,-1]
Test <- dat_raw[-index,-1]
```

```{r}
#implement random forest
library(randomForest)
library(caret)
learn_rf <- randomForest(diagnosis~., data=Train, ntree=500, proximity=T, importance=T)
#variance importance plot
varImpPlot(learn_rf)

#append predicted values onto test set to plot confusion matrix
Test$predicted <- round(predict(learn_rf ,Test),0)

plotConfusionMatrix <- function(testset, sSubtitle) {
    tst <- data.frame(testset$predicted, testset$diagnosis)
    opts <- c("Predicted", "True")
    names(tst) <- opts
    cf <- plyr::count(tst)
    cf[opts][cf[opts]==0] <- "Benign"
    cf[opts][cf[opts]==1] <- "Malignant"
    ggplot(data =  cf, mapping = aes(x = True, y = Predicted)) +
      labs(title = "Confusion matrix", subtitle = sSubtitle) +
      geom_tile(aes(fill = freq), colour = "grey") +
      geom_text(aes(label = sprintf("%1.0f", freq)), vjust = 1) +
      scale_fill_gradient(low = "gold", high = "tomato") +
      theme_bw() + theme(legend.position = "none")
}

#calculate and plot AUC
library(pROC)
print(sprintf("Area under curve (AUC) : %.3f",auc(Test$diagnosis, Test$predicted)))




plotConfusionMatrix(Test,"Prediction using RandomForest with 500 trees")
#check prediction accuracy
```

##Refine covariates according to the result of the varimpplot
```{r}
features_list <-  c("perimeter_worst", "area_worst", "concave.points_worst", "radius_worst", "concavity_mean", "concavity_worst","area_se", "concave.points_mean","texture_worst", "area_mean", "texture_mean", "area_mean", 
 "radius_mean", "radius_se", "perimeter_mean", "perimeter_se", "compactness_worst", "smoothness_worst", "concavity_se", "fractal_dimension_worst", "symmetry_worst",  "diagnosis")



#define train and validation set

Train_red = Train[,features_list]

Test_red =  Test[,features_list]

#training
train_red_rf <- randomForest(diagnosis ~.,Train_red,ntree=500,importance=T)
#get prediction
Test_red$predicted <- round(predict(train_red_rf ,Test_red),0)

plotConfusionMatrix(Test_red,"Prediction using RandomForest with reduced features set")

#increase in true negative
#calculate AUC
print(sprintf("Area under curve (AUC) : %.3f",auc(Test_red$diagnosis, Test_red$predicted)))

#reducing the features only increased prediction rate by 0.007
