---
title: "R Notebook"
output: html_notebook
---

# About the notebook
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(caret)
library(class)        ### for knn
library(glmnet)       ### for LASSO/Ridge
library(randomForest) ### for rf
library(gridExtra)
```

First I will write the cross validation for each methods

# Import data

read in data
```{r, message = FALSE, warning = FALSE}
dat_raw = read_csv("./data/data.csv")
dat_raw = dat_raw[, 1:32]
colnames(dat_raw) = str_replace(
    string = colnames(dat_raw), 
    pattern = " ", 
    replacement = "_")

dat_raw$diagnosis = factor(dat_raw$diagnosis, levels = c("B", "M"))
dat_raw$label     = as.integer(dat_raw$diagnosis) - 1
```

check if the label is balanced
```{r}
table(dat_raw$diagnosis)
```

upsampleing to make the label balance
```{r}
### initialization
set.seed(123)

### upsampleing of malignant tumor
numdiff = table(dat_raw$diagnosis)["B"] - table(dat_raw$diagnosis)["M"]
dat = dat_raw %>% filter(diagnosis == "M") %>% sample_n(size = numdiff) 
dat = bind_rows(dat_raw, dat)

### check if the label is balanced
table(dat$diagnosis)
```


preprocess the data
```{r}
dat_prep  = dat  %>% select(-id, -diagnosis)
dat_covar = dat_prep %>% select(-label)
dat_label = dat_prep %>% select( label)
```



# cross validation of random Forest 

fit a model with a set of parameter to make sure the code for model fitting works properly
```{r}
### parameters
mtry   = 5
ndsize = 5

### arrange the data
dat       = dat_covar %>% scale %>% as.data.frame
dat$label = dat_label$label

### fit random forest
rfFit <- randomForest(
    factor(label) ~ ., data = dat, 
    do.trace = 100, importance = T, ntree = 10,
    mtry     = mtry, 
    nodesize = ndsize)

### 
```


```{r}
### intialization 
MTRY_ALL     = c(1, 10, 20, 30) #seq(1, 30, by = 10)
NODESIZE_ALL = c(5, 10, 15, 20)

### combination of mtry and nodesize
params = expand.grid(mtry = MTRY_ALL, nodesize = NODESIZE_ALL)
params$error = NA
    
### fit random forest
rf_all = list()
for (idx in seq_len(nrow(params))){
    param  = params[idx, ]
    mtry   = param$mtry
    ndsize = param$nodesize
    
    rf <- randomForest(
            factor(Genotype) ~ ., 
            data       = model_mat, 
            importance = T, 
            mtry       = mtry,
            nodesize   = ndsize,
            ntree      = 200)
    
    rf_all[[idx]] = rf
    params$error[idx] = mean((model_mat$Genotype - rf$votes[, 2])^2)
} # end for loop
```

To-Do list
# Cross validation of k nearest neighbor
# Cross validation of logistic regression
# Cross validation of all methods together

